---
title: "Homework1_G15"
authors: "Adriano Fragomeni, Riccardo Gobbo, Ufuk Caliskan"
output: html_document
runtime: shiny
---

##Import Libraries
```{r message=FALSE, warning=FALSE}
library(ggmap)
library(gplots)
library(plyr)
library(pracma)
library(shiny)
library(Matrix)
library(ks)
library(sparr)
require(data.table)
library(MASS)
library(meanShiftR)
library(plotly)
source("Function.R")
```

The goal of this project is to study data coming from a tracking device where there are 60 tracks:
the first part of this project is to try to estimate the density with the gaussian KDE and to use mean-shift
algorithm to cluster points.
The second part is to consider each point as an element of a track and to find the top-5 paths and the bottom-5
paths, using the Boxcar Kernel estimation of the Hausdorff distances between each track.

##LOAD DATA

The original dataset is based on 5 columns:
1)longitude
2)latitude
3)elevation
4)time
5)run-id

For the following analysis the relevant variable are just three:
1)longitude
2)latitude
3)time 

```{r}
# load data 
load("trackme.RData")
df <- as.data.frame(cbind(runtrack$lon,runtrack$lat, runtrack$time))
colnames(df) = c("lon", "lat", "DT")
set<-split(df,runtrack$id)
# number of tracks
n<-length(set)  
```

The dataset is large, around $60000$ points, beacause the data comes from a tracking device, and so it is
possible that there is noise, so to remove it, the "noise" function was used:
for each run it removes the rows where the time difference is smaller than 5 seconds

```{r}
#Clean data
set.cleaned<-lapply(set, noise)
subtrack <- ldply(set.cleaned, data.frame)
subtrack <- subtrack[,c(2,3)]
set<-split(df,runtrack$id)
```

After cleaning data, the new dataset has around 30000 observations; the next step is to visualize them to have a
general idea on how they are allocated on the Rome map

```{r message=FALSE, warning=FALSE}
# Map boundaries 
location<- c(min(subtrack$lon, na.rm = T), min(subtrack$lat, na.rm = T), 
             max(subtrack$lon, na.rm = T), max(subtrack$lat, na.rm = T))
# Get the map from Google
myMapInD <- get_map(location = location, maptype = "roadmap", zoom = 13)
# Plot gps coordinates 
gp<-plot.map(myMapInD,subtrack,.5,I("red"),.03,'points on Rome map')

print(gp)
```

Some qualitative observations can be done by looking directly this plot,:
The points are more concentrated in the range of $latitude=(12.500-12.550)$ and $longitude=(41.90-41.92)$, so it
is likely that there will be some peaks of the KDE in this area.
Let us zoom in this area to show better what happens there:

```{r warning=FALSE}
# Zoom in
zoom<- c(12.500, 41.90, 12.550, 41.92) 
# Get the map from Google and plot 
reMapInD <- get_map(location = zoom, maptype = "roadmap") 
gp<-plot.map(reMapInD,subtrack,1,I("red"),.05,'Zoom of more concetrated points')
print(gp)

```

#FIRST PART

##CHOOSING BANDWIDTH

The bandwidths are the crucial selected parameters which are necessary to obtain a good estimation of density; if
the selected bandwidths are too small, the estimation will be rough (undersmoothing), while larger bandwidths
give large smoother estimates (oversmoothing).
Since we are going to use not scaled data, the bandwidth is calculated for each dimension ('lat','lon')

```{r}
#bandwidth
h <- c(bw.nrd(subtrack$lon),bw.nrd(subtrack$lat))
h
```

The function above uses a rule of thumb to calculate the bandwidth of data, using the Scott factor:
$n^{\frac{-1}{d+4}}$
where $n$ is the number of data points and $d$ is the number of dimensions.

##KDE

After having fixed 'h', we can calculate the Gaussian kernel function to estimate the density of data, in
contrast to the bandwidth the choice of the shape of the Kernel does not influence the KDE.
In this project the Gaussian kernel is chosen:

$K(x)=\frac{1}{\sqrt{2\pi}}\times{\exp(-\frac{x^2}{2})}$

```{r}
#Kernel density estimation
kd <- kde2d(subtrack$lon, subtrack$lat, h = h)
#plot kernel
p3 <- plot_ly(x = kd$x, y = kd$y, z = kd$z)%>% add_surface()
p3
```

From the plot, it seems that there can be at least four clusters:
three of them are associated to the three highest peaks of the estimated density and the other one has points
with lower density.

##MEAN SHIFT

The mean-shift algorithm is used to cluster data, as following:
Given a dataset $X=\{X_1,...,X_n\} \quad iid\quad p$, we construct an estimate $\hat{p}$ of the density.
Let $\{\hat{m_1},..,\hat{m_k}\}$ be the estimated modes, and let$\{\hat{A}_k\}_{k=1}^K$ be the corresponding
ascending sets of points implied by $\hat{p}$.
The empirical clusters $\{C_1,...,C_k\}$ are defined as
$C_k=\{X_i:X_i \in \hat{A_k}), \forall k \in \{1,..,K\}$
To do this, the 'meannShift' function of the package 'meanshiftR' was used, although data are cleaned, each point
is really close to each other in terms of coordinates, so if the mean-shift algorithm runs without parameters,
apart from the 'h', there will be just one cluster

```{r}
# mean-shift algorithm
ms1 <- meanShift(as.matrix(subtrack), bandwidth = h)
```

and one mode, which corresponds to the highest peak of the KDE

```{r}
# modes of cluster
modes1<-as.data.frame(unique(ms1$value))
colnames(modes1)<-c('lon','lat')

```

it means that each points are attracted by this peak and so they are going to set just one cluster, as you can
see from the following plot

```{r}
# plot cluster
gp<-plot.map(myMapInD,subtrack,.4,ms1$assignment,.08,'Bad mean-shift')+geom_point(data=modes1,aes(x=lon,y=lat),col='brown',size=3,shape=18)

print(gp)
```

but this is not good, it means that we put all points in one set and so there is not any relevant information, so
the 'epsilonCluster' parameter was introduced to avoid this problem, indeed it is the minimum distance between
distinct clusters, the chosen value is equal to $1.8e-7$ because points are very close to each other, for this
reason the value is small.
Run the mean-shift, using this new parameter

```{r}
# mean-shift algorithm
ms <- meanShift(as.matrix(subtrack), bandwidth = h, epsilonCluster = 1.8e-7)
```

The number of the cluster are 6 and the coordinates of these modes are the following one

```{r}
# modes of cluster
modes<-as.data.frame(unique(ms$value))
colnames(modes)<-c('lon','lat')
```

Take a look at the cluster, plotting them on the Rome map

```{r}
# plot cluster
gp<-plot.map(myMapInD,subtrack,.4,ms$assignment,.08,'Good mean-shift')+geom_point(data=modes,aes(x=lon,y=lat),col='brown',size=3,shape=18)

print(gp)
```

This cluster is completely different from the previous one, the reason is that the 'epsilonCluster' parameter
allows to cluster data; as you can see, the area with the lowest density on the left of the plot is divided in
three clusters, this means that the algorithm finds three relevant peaks instead of the qualitativie analysis
done by looking the KDE plot, whereas on the right of the plot, where there is a majority concentration of points
there are also three clusters, probably associated to the peaks that it is possible to see by looking at the plot
of the KDE, as it was said before.


#SECOND PART

Actually each point is not a single point in a plane, but it is a part of a track, a single curve in the plane.

##HAUSDORFF DISTANCE

To identify which tracks belong to the top-5 and bottom-5, the first thing to do is to create a matrix 60x60,
where 60 is the number of the total tracks, with the distance of each track to the others, to do this the
Hausdorff metric is used.
In this metric it is possible to define the Hausdorff distance
Given 2 sets of points $A$ and $B$:

$dist_H(A,B)=max\{sup(x \in A)inf(y\in B)dist(x,y),sup(x \in B)inf(y\in A)dist(x,y)\}$

How does the Hausdorff distance work?
Given a curve $\gamma$, it is possible to calculate $dist_H(\gamma,G_i)$, it depends only on the distance
$dist(x,y)$ between each point in $\gamma$ and $G_i$, which is calculated using the Euclidean distance.

```{r}
# Hausdorff matrix
mat<-haud_dist(set.cleaned,n)
rowscols <- which(upper.tri(mat), arr.ind=TRUE)
mat<-sparseMatrix(i=rowscols[,1],    #rows to fill in
             j=rowscols[,2],    #cols to fill in
             x=mat[upper.tri(mat)], #values to use 
             symmetric=TRUE,    #make it symmetric
             dims=c(nrow(mat),nrow(mat))) #dimensions
```

##TOP-5 AND BOTTOM-5 TRACKS

Also in this situation the estimation of the curve's density is relevant to find the top-5 and the bottom-5, in
this case we are going to estimate it using the unormalized boxcar-Kernel estimator:

$\hat{q_{\epsilon}(\gamma)=\frac{1}{n}\displaystyle\sum_{i=1}^{n}I(dist_H(\gamma,G_i)\leq\epsilon)}$

As you can see from the above formula, the choice of $\epsilon$ is important to explore and analyze tracks, so
the best idea is to explore data for different values of $\epsilon$ and see what happens

```{r}
# behaviour of kernel by epsilon
inputPanel(
  sliderInput("epsilon", label = "epsilon", min = 0.001, value = 0.011, max = 0.1, step = 0.0001))

q.kern <- reactive({q_density(mat, input$epsilon,n)})

renderPlot({
  plot.kern(q.kern())
})

```

The plot above shows the behaviour of curve's density, it changes by varying $\epsilon$, if it is too small, for
example 0.002 or 0.003, it is possible to see that we have a straight line close to zero ( it means that many
curves have a small density) and just a few of them with an higher one; increasing $\epsilon$ the density tends
to vary a lot, whereas if the $\epsilon$ values is too large, around 0.062, there is an opposite situation of the
small $\epsilon$, in which the majority of curves have a maximum density (1) and a few curves with lower one,
when the $\epsilon$ is larger than 0.063 all curves have a density equal to 1,as it is possible to imagine if you
look at the formula of boxcar-kernel estimator.
The following analysis is based on just 4 values of $\epsilon$, one of them is chosen by looking at the plot
(0.0115) and the other three using the quantiles of the Hausdorff distances (25%,50%,75%) 

```{r message=FALSE, warning=FALSE, echo=FALSE}
#quantile of Hausdorff distance
eps<-unname(round(quantile(mat,probs=c(0.25,0.5,0.75)),4))
eps<-c(eps,0.0110)
```

Now plot the top-5 and bottom-5 paths for these four values of $\epsilon$

```{r}
# plot
inputPanel(
  selectInput("eps", label = "eps", choices = eps,selected = eps[4]))

q.kern1 <- reactive({q_density(mat, input$eps,n)})

#min tracks
renderTable({min.find(q.kern1())})

renderPlot({
  plot.path(myMapInD,df.min(runtrack,set.cleaned,min.find(q.kern1())),1.5,.6,'Bottom-5 tracks')})

#max tracks
renderTable({max.find(q.kern1())})

renderPlot({
plot.path(myMapInD,df.max(runtrack,set.cleaned,max.find(q.kern1())),1.5,.6,'Top-5 tracks')})

```

The value of chosen $\epsilon$, looking at the data, is equal to the first quantile of Hausdorff distance, so the
result are completely the same.
It is possible to see that for each value of $\epsilon$, taken into account, the runtrack 2 and 3 are always in
the bottom-5, so it means that they are the less frequent tracks that you (Pierpaolo) run.
The top-5 tracks are often in the same area of Rome (close to Sapienza University and via Tiburtina), so probably
this is your favorite area where you go to have fun, just the third quantile shows that you love Villa
Borghese....you know...it is a park...amazing place!!!

#CONCLUSION
From all this information we can conclude that you live, probably, in this area

```{r}
zoom<- c(12.530, 41.90, 12.550, 41.92) 
# Get the map from Google (default) and plot 
# Get the map from Google and plot 
reMapInD <- get_map(location = zoom, maptype = "roadmap") 
gp<-plot.map(reMapInD,subtrack,1,I("red"),.05,'Your Home')
print(gp)

```


